#CodeBook

The UCI HAR dataset is an experimental dataset.The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz were captured. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

For each record it is provided:

- Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration.
- Triaxial Angular velocity from the gyroscope. 
- A 561-feature vector with time and frequency domain variables. 
- Its activity label. 
- An identifier of the subject who carried out the experiment.

The following files are available for the train and test data

- 'train/X_train.txt': Training set.

- 'train/y_train.txt': Training labels.

- 'test/X_test.txt': Test set.

- 'test/y_test.txt': Test labels.

#Project Code Discription

First load the necessary libraries in r

```{r,eval=FALSE}
library(data.table)
library(plyr)
library(dplyr)
```
### *Directories and file description*

All the files stores in UCI HAR Dataset directory.

```{r}
uci_data<- "UCI\ HAR\ Dataset"
feature_file<-paste(uci_data,"/features.txt",sep="")
activity_file<-paste(uci_data,"/activity_labels.txt",sep="")
subject_file_t<-paste(uci_data,"/train/subject_train.txt",sep="")
y_file_t<-paste(uci_data,"/train/y_train.txt",sep="")
X_file_t<-paste(uci_data,"/train/X_train.txt",sep="")
subject_file_test<-paste(uci_data,"/test/subject_test.txt",sep="")
y_file_test<-paste(uci_data,"/test/y_test.txt",sep="")
X_file_test<-paste(uci_data,"/test/X_test.txt",sep="")
```

Train and Test files read by following r commands:
```{r}
activitylabels<-read.table(activity_file)
subject_train<-read.table(subject_file_t,col.names=c("subject"))
y_train<-read.table(y_file_t,col.names=c("activity"))
X_train<-read.table(X_file_t,nrows=7352, comment.char="")
subject_test<-read.table(subject_file_test,col.names=c("subject"))
y_test<-read.table(y_file_test,col.names=c("activity"))
X_test<-read.table(X_file_test, nrows=2947, comment.char="")
```

### Step1: Merge the train and test set to create one dataset
The data in X_train, y_train and subject_train is combined using cbind and stored in training_data. 
Similarly test data of all three files (X_test,y_test,subject_test) is combined and stored in testing_data. Finally, training_data and testing_data is combined row-wise using rbind command and merged data is stored in merge_data

```{r}
training_data<-cbind(X_train,subject_train,y_train)
testing_data<-cbind(X_test,subject_test,y_test)
merge_data<-rbind(training_data,testing_data)
```

### Step2: Extracts only the measurements on the mean and standard deviation for each measurement.

read the feature list from features.txt with columnames=("id","names"). select the column indexes with 
mean or std or subject or activity.Data is extracted for selected ids.
```{r}
feature_list <- read.table(feature_file, col.names = c("id", "name"))
features <- c(as.vector(feature_list[, "name"]), "subject", "activity")
selected_feature_ids <- grepl("mean|std|subject|activity", features) & !grepl("meanFreq", features)
extracted_data = merge_data[, selected_feature_ids]
```

### Step3:Uses descriptive activity names to name the activities in the data set
The data is classified based on 6 activities, which is described in activity_labels.txt and can be 
load using R command:

```{r}
activities <- read.table(activity_file, col.names=c("id", "name"))
activities
```

```{r}

##   id               name
## 1  1            WALKING
## 2  2   WALKING_UPSTAIRS
## 3  3 WALKING_DOWNSTAIRS
## 4  4            SITTING
## 5  5           STANDING
## 6  6             LAYING
```
descriptive name to the activity in the data set. there are 6 activities. 

```{r}
for (i in 1:nrow(activities)) {
  extracted_data$activity[extracted_data$activity == activities[i, "id"]] <- as.character(activities[i, "name"])
}
```

### Step 4:Appropriately labels the data set with descriptive variable names. 

In order to make variable more understandable and readable, 
the following acronyms can be replaced:

   - "Acc"" can be replaced with Accelerometer

   - "Gyro" can be replaced with Gyroscope

   - "BodyBody" can be replaced with Body

   - "Mag" can be replaced with Magnitude

   - "Character" f can be replaced with Frequency

   - "Character" t can be replaced with Time

```{r}
selected_feature_names <- features[selected_feature_ids]
selected_feature_names <- gsub("\\(\\)", "", selected_feature_names)
selected_feature_names <- gsub("Acc", "-acceleration", selected_feature_names)
selected_feature_names <- gsub("Mag", "-Magnitude", selected_feature_names)
selected_feature_names <- gsub("^t(.*)$", "\\1-time", selected_feature_names)
selected_feature_names <- gsub("^f(.*)$", "\\1-frequency", selected_feature_names)
selected_feature_names <- gsub("(Jerk|Gyro)", "-\\1", selected_feature_names)
selected_feature_names <- gsub("BodyBody", "Body", selected_feature_names)
selected_feature_names <- tolower(selected_feature_names)
names(extracted_data)<- selected_feature_names
names(extracted_data)
```
```{r}
##  [1] "body-acceleration-mean-x-time"                  
##  [2] "body-acceleration-mean-y-time"                  
##  [3] "body-acceleration-mean-z-time"                  
##  [4] "body-acceleration-std-x-time"                   
##  [5] "body-acceleration-std-y-time"                   
##  [6] "body-acceleration-std-z-time"                   
##  [7] "gravity-acceleration-mean-x-time"               
##  [8] "gravity-acceleration-mean-y-time"               
##  [9] "gravity-acceleration-mean-z-time"               
## [10] "gravity-acceleration-std-x-time"                
## [11] "gravity-acceleration-std-y-time"                
## [12] "gravity-acceleration-std-z-time"                
## [13] "body-acceleration-jerk-mean-x-time"             
## [14] "body-acceleration-jerk-mean-y-time"             
## [15] "body-acceleration-jerk-mean-z-time"             
## [16] "body-acceleration-jerk-std-x-time"              
## [17] "body-acceleration-jerk-std-y-time"              
## [18] "body-acceleration-jerk-std-z-time"              
## [19] "body-gyro-mean-x-time"                          
## [20] "body-gyro-mean-y-time"                          
## [21] "body-gyro-mean-z-time"                          
## [22] "body-gyro-std-x-time"                           
## [23] "body-gyro-std-y-time"                           
## [24] "body-gyro-std-z-time"                           
## [25] "body-gyro-jerk-mean-x-time"                     
## [26] "body-gyro-jerk-mean-y-time"                     
## [27] "body-gyro-jerk-mean-z-time"                     
## [28] "body-gyro-jerk-std-x-time"                      
## [29] "body-gyro-jerk-std-y-time"                      
## [30] "body-gyro-jerk-std-z-time"                      
## [31] "body-acceleration-magnitude-mean-time"          
## [32] "body-acceleration-magnitude-std-time"           
## [33] "gravity-acceleration-magnitude-mean-time"       
## [34] "gravity-acceleration-magnitude-std-time"        
## [35] "body-acceleration-jerk-magnitude-mean-time"     
## [36] "body-acceleration-jerk-magnitude-std-time"      
## [37] "body-gyro-magnitude-mean-time"                  
## [38] "body-gyro-magnitude-std-time"                   
## [39] "body-gyro-jerk-magnitude-mean-time"             
## [40] "body-gyro-jerk-magnitude-std-time"              
## [41] "body-acceleration-mean-x-frequency"             
## [42] "body-acceleration-mean-y-frequency"             
## [43] "body-acceleration-mean-z-frequency"             
## [44] "body-acceleration-std-x-frequency"              
## [45] "body-acceleration-std-y-frequency"              
## [46] "body-acceleration-std-z-frequency"              
## [47] "body-acceleration-jerk-mean-x-frequency"        
## [48] "body-acceleration-jerk-mean-y-frequency"        
## [49] "body-acceleration-jerk-mean-z-frequency"        
## [50] "body-acceleration-jerk-std-x-frequency"         
## [51] "body-acceleration-jerk-std-y-frequency"         
## [52] "body-acceleration-jerk-std-z-frequency"         
## [53] "body-gyro-mean-x-frequency"                     
## [54] "body-gyro-mean-y-frequency"                     
## [55] "body-gyro-mean-z-frequency"                     
## [56] "body-gyro-std-x-frequency"                      
## [57] "body-gyro-std-y-frequency"                      
## [58] "body-gyro-std-z-frequency"                      
## [59] "body-acceleration-magnitude-mean-frequency"     
## [60] "body-acceleration-magnitude-std-frequency"      
## [61] "body-acceleration-jerk-magnitude-mean-frequency"
## [62] "body-acceleration-jerk-magnitude-std-frequency" 
## [63] "body-gyro-magnitude-mean-frequency"             
## [64] "body-gyro-magnitude-std-frequency"              
## [65] "body-gyro-jerk-magnitude-mean-frequency"        
## [66] "body-gyro-jerk-magnitude-std-frequency"         
## [67] "subject"                                        
## [68] "activity"
```

###Step5:From the data set in step 4, creates a second, independent tidy data set with the average 
###of each variable for each activity and each subject.
created tidy Data as a data set with average for each activity and subject

```{r,eval=FALSE}
selected_tidy_data <- ddply(extracted_data,c("subject","activity"),numcolwise(mean))
write.table(selected_tidy_data,file="selected_tidy_data.txt")
```



